Statistical Mechanics Assignment One

### 2016-04-15 stat_mech/latex/a1.tex
blablabl

>>>>

\section{Problem Two: Sound compression of a liquid}
We use the following variable scheme: $\Delta x$ is the initial length of our 1D
liquid. $\epsilon$ is the change in our 1D liquid length due to rarefaction
on the left side. $\epsilon + \Delta \epsilon$ is the change in length to the right
side.

\begin{align*}
  K_S &= -\frac{1}{V}\left(\frac{dV}{dP}\right)_S
  = \frac{-1}{V}\frac{\Delta V}{\Delta P}& \hspace{2cm}&\mbox{P-P0: displacement from eq pressure}\\
  &= \frac{-1}{A\Delta x}\frac{A\Delta\epsilon}{P-P_0}& \hspace{2cm}&\mbox{A: Area}\\
  \Delta P &= -\frac{1}{K_s}\frac{d\epsilon}{dx}\\
\end{align*}

We then need to find the time derivative of $\epsilon$, which is the force on our
1D cell, or how fast it is growing. The pressure wave is moving in the negative
x direction, so the force on the right of the cell is $P_0 + \Delta P$. On the
left, our cell is being pushed by the normal pressure $P_0$. Thus by Newton's First
Law we see that:\\

\begin{align*}
  m\left(\frac{d^2\epsilon}{dt^2}\right) &= P_0A - (P_0+\Delta P)A = -\Delta PA\\
  \rho A \Delta x \frac{d^2\epsilon}{dt^2} &= -A\frac{dP}{dx}\Delta x\\
  \rho \frac{d^2\epsilon}{dt^2} &= -\frac{dP}{dx}\\
  \rho \frac{d^2\epsilon}{dt^2} &= -\frac{d \Delta P}{dx}\\
  \rho \frac{d^2}{dt^2}\frac{d\epsilon}{dx} &= -\frac{d^2P}{dx^2}\\
  \rho \frac{d^2}{dt^2}\left(-PK_s\right) &= -\frac{d^2P}{dx^2}\\
  \rho K_s \frac{d^2P}{dt^2}\left(P\right) &= \frac{d^2P}{dx^2}\\
  \frac{d^2P}{dt^2} &= \frac{1}{\rho K_s }\frac{d^2P}{dx^2}\\
\end{align*}

Solution inspired by several people from SPS.\\

\section{Problem Three: Solid $\Delta T, \Delta P$ relation}

\begin{align*}
  dS &= \left(\frac{dS}{dT}_P\right)dT + \left(\frac{dS}{dP}_T\right)dP = 0\\
  \left(\frac{dS}{dT}_P\right)dT &= -\left(\frac{dS}{dP}_T\right)dP\\
  \frac{C_P}{T}dT &= \left(\frac{dV}{dT}_P\right)dP\\
  \frac{C_P}{T}dT &= V\alpha dP\\
  \frac{C_P}{TV\alpha} &= -\left(\frac{dP}{dT}_S\right)\\
  \frac{C_P}{TV\alpha} &= \frac{\Delta P}{\Delta T}\\
  \Delta T &= \frac{TV\alpha}{C_P} = \frac{\alpha T}{\rho c_p} \hspace{2cm}
  c_p \mbox{ is the volumetric specific heat}
\end{align*}
